### 2020-10-08 camila/report/concl.tex
first complete paper draft

\section{Concluding Remarks}
\label{sec.concl}

This manuscript provides a detailed description of a network-based analysis
methodology for the discovery of key genes responding to
a specific treatment in an organism. It links transcriptomic with
phenotypic data, and identifies overlaping gene modules.

The proposed methodology is inspired by the workflow suggested in the
WGCNA methodology~\cite{langfelder2008wgcna}. The main steps are the
preprocessing of the gene expression data, the construction of a
co-expression network, the detection of modules within the network,
the relation of modules with external information (e.g. phenotypic data)
and the enrichment of the identified key genes with additional information.
The WGCNA methodology and therefore the proposal as well, are structured
in a modular way, which allows modifying and  exploring different techniques
in each step of the process.

The proposed approach is designed to integrate expression data
measured under two different conditions (namely, control and treatment),
unlike the usually co-expression-based approaches which work with both
conditions independently or consider only a single condition. For this
purpose, an approach similar to that proposed in~\cite{du2019network}
is used, where the control and treatment data are compiled in a single
matrix using the log fold change measure. Thus, the input to construct
the co-expression network is not the expression data, but instead the changes
in the expression levels from one condition to the other, making room for capturing
the signal of changes caused by the treatment.

Another important feature in the proposed methodology is the module
detection technique. The co-expression network is computed as in
WGCNA until a scale-free network is obtained. This network is then
used to apply the HLC algorithm, a clustering technique capable of
detecting overlapping communities. Several approaches of module
detection from gene expression have been proposed and were evaluated
in~\cite{saelens2018comprehensive}. Most of them focuses only on
disjoint (non-overlapping) communities, and the described techniques
dealing with overlaps are not clustering but biclustering and
decomposition methods. It is well known that communities in real
networks are overlapping~\cite{palla2005uncovering}. Thus, the approach
presented in this work can be seen as a generalization of the previous 
approaches such as WGCNA.

The proposed methodology was applied in a case study with rice under salt
stress. The results show a group of 14 genes in which $2$ of them are
related to the response to saline stress according to previous
studies, validating the ability of the method to detect this kind of
key genes. As future work, other overlapping module detection and selection
techniques should be used instead HLC and LASSO, respectively. The
combination of these techniques would allow finding target genes for
future biological studies that evaluate their potential as genes that
respond to salt stress in rice. Also, laboratory experimentation needs to be
conducted on via to verify the findings of this paper in relation to salinity 
stress. Finally, this study can be extended to other stresses and even to other 
crops.