### 2020-10-08 camila/report/prelim.tex

\section{Preliminaries}
\label{sec.prelim}

This section presents preliminaries on networks, hierarchical link
clustering and the LASSO linear regression technique.

\subsection{Co-expression network}

A network is an undirected graph $G=(V,E)$ where
${V=\{v_1,v_2,\ldots,v_{n}\}}$ is a set of \textit{vertices} or
\textit{nodes} and ${E=\{e_1,e_2,\ldots,e_q\}}$ is a set of
\textit{edges} or \textit{links} that connect vertices. In a gene
co-expression network, each node corresponds to a gene. A pair of
genes is connected if they show similar expression
patterns. A simple and unweighted network can be represented by an
adjacency matrix $A \in \{0,1\}^{n \times n}$ that is symmetric with a
positive one in the positions $(v_i,v_j)$ and $(v_j,v_i)$ whenever
there is an edge connecting vertices $v_i$ and $v_j$, and zeros
elsewhere. Co-expression networks are of biological interest because
the co-expressed genes are usually controlled by the same
transcriptional regulatory pathway, functionally related or members of
the same pathway or metabolic complex.

\subsection{Hierarchical Link Clustering}

The Hierarchical Link Clustering (HLC) algorithm was proposed by Ahn
et al.~\cite{ahn2010link}. The HLC approach represents communities as
groups of links (rather than nodes), and each node inherits all
memberships of its links and can thus belong to multiple, overlapping
communities. It maps links to nodes and connects them if a pair of
links shares a node. The similarity between two links $e_{ik}$ and
$e_{jk}$ is computed using the Jaccard index
%
\begin{equation}\label{eq:jaccard}
  S(e_{ik},e_{jk}) = \frac{\vert n(i) \cap n(j) \vert}{\vert n(i) \cup n(j) \vert},
\end{equation}
%
where $n(i)$ denotes the set containing exactly node $i$ and its
neighbors. The algorithm uses single-linkage hierarchical clustering
to build a dendrogram in which each leaf is a link from the original
network and branches represent link communities. Hierarchical
clustering algorithms repeatedly merge groups until all elements are
members of a single cluster. For the purpose of finding meaningful
communities, it is crucial to know where to partition the
dendrogram. In this case, the most relevant communities are
established at the maximal partition density $D$, a function based on
link density inside communities measuring the quality of a link
partition. The partition density $D$ has a single global maximum along
the dendrogram in almost all cases, because its value is the average
density at the top of the dendrogram (a single giant community with
every link and node) and it is very small at the bottom of the
dendrogram (most communities consists of a single link). In
particular, it is the case that $D = 1$ when every community is a
fully connected clique and $D = 0$ when each community is a tree. If a
community is less dense than a tree (i.e. when the community subgraph
has disconnected components), then such a community contribute
negatively to $D$, which can take negative values. The minimum density
inside a community is $-2/3$, given by one community of two
disconnected edges. Since $D$ is the average of the intra-community
density, there is a lower bound of $-2/3$ for $D$. Computing $D$ at
each level of the link dendrogram can help the purpose of picking the
best level to cut, although meaningful structure could exist above or
below the threshold. The output of cutting is a set of node clusters,
where each node can participate in multiple communities.

\subsection{Least Absolute Shrinkage Selector Operator}

The Least Absolute Shrinkage Selector Operator (LASSO) is a
regularized linear regression technique. It combines a regression
model with a procedure of contraction of some parameters towards zero
and selection of variables, imposing a restriction or a penalty on the
regression coefficients. In other words, LASSO solves the least
squares problem with restriction on the $ L_1$-norm of the coefficient
vector. It can be especially useful to solve problems where the number
of variables (e.g., genes) $ n $ is much greater than the number of
samples $ m $ (i.e., $ n \gg m $).

Consider a dataset consisting of $m$ samples, each of which consists
of $n$ covariates and a single outcome. Let $y_i$ be the outcome and
$x_i := (x_1,...,x_n)$ be the covariate vector for the $i$-th
sample. The objective of LASSO is to solve
%
\begin{equation}
\min \left\lbrace\sum_{i=1}^{m}{\left( y_i-\sum_{j=1}^n{\beta_j
    x_{ij}}\right)^2} \right\rbrace \quad , \quad \textrm{subject to}
\quad \sum_{j=1}^n\abs{\beta_j}\leq s.
\end{equation}

Equivalently, in the Lagrangian form, it minimizes

\begin{equation}
  \sum_{i=1}^{p}{\left( y_i-\sum_{j=1}^n{\beta_j x_{ij}}\right)^2} +
  \lambda \sum_{j=1}^n\abs{\beta_j}
\end{equation}
%
where $s$ is the regularization penalty and $\lambda \geq 0$ is the
corresponding Lagrange multiplier. Since the $\lambda$ value
determines the degree of penalty, the accuracy of the model depends on
its choice. Cross-validation is often used to select the
regularization parameter, choosing the one that minimizes the
mean-squared error.