### 2020-10-08 camila/report/main.tex

\documentclass[runningheads]{llncs}

\input{packages}
\input{macros}

\begin{document}
%
\title{Using Complex Networks for Identifying Potentially New Saline
  Stress Responsive Genes in Rice}
%
\author{Camila Riccio \and Jorge Finke \and Camilo Rocha}
%% \author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
%% Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
%% Third Author\inst{3}\orcidID{2222--3333-4444-5555}}

\authorrunning{C. Riccio et al.}

\institute{Pontificia Universidad Javeriana, Cali, Colombia
  %% \email{lncs@springer.com}
}  
%
\maketitle

\input{abstract}

\input{intro}

\input{prelim}

\input{framework}

\input{case}

\input{concl}

\bibliographystyle{splncs04}
\bibliography{biblio}

\end{document}