### 2020-10-15 nicolas/PPIPL3/ReportV4.tex
Last update on report

%% LyX 2.3.4.2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{verbatim}
\usepackage{float}
\usepackage{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

\makeatother

\usepackage{babel}
\begin{document}
\title{Prediction of Protein-Protein Interactions on the Human and Rice Interactome }
\author{Nicolás Antonio López Rozo}
\maketitle
\begin{abstract}

Network analyses for predicting unmapped protein-protein interactions
(PPI) suggest that the higher the number of paths of length 3 (L3)
between two proteins, the more likely they are to interact. This paper
extends previous work based on the L3 principle by taking into account
the representation learning of node features of the PPI network. In
particular, we train an \texttt{XGBoost} model using L3 and handcrafted
features, as well as embeddings from \texttt{Node2Vec}. Our main result
shows that while L3 is an important feature for predicting links,
best performance is achieved when combined with edge embeddings. The
proposed approach is evaluated for the human and rice interactomes.
\end{abstract}
\begin{comment}
need include a feature that compares the dimension of the embedding
is useful for prediction with L3 to show that they independent measures
- we do not want to be learning L3 through node2vec
\end{comment}

\section{Introduction}

\begin{comment}
Need to add other relevant references
\end{comment}

Proteins are the key actors of biological processes inside cells.
As they carry out a variety of tasks, proteins do not work as single
agents but rather as a part of dynamic networks of protein-protein
interactions (PPI) \cite{Lin2017}. PPI networks play a key role in
a variety of interdependent mechanisms, including signal transduction,
homeostasis control, stress responses, plant defense and organ formation.
At the molecular level, PPIs play an essential role in many physiological
and developmental processes, including protein phosphorylation, transcriptional
co-factor recruitment and transporter activation \cite{Zhang_2010_PPI}.

Prediction of potentially relevant yet unexplored PPIs is a current
research topic on bioinformatics. Several authors have proposed different
methods for extrapolating information from the existing PPI networks.
Kovacs et al. (2019) introduces an approach which relies on counting
paths of length 3 (L3) to predict interactions among proteins for
a variety of model organisms. The proposed approach outperform previous
methods in different PPI networks of yeast (\emph{S. cerevisiae}),
Arabidopsis (\emph{A. thaliana}), worm (\emph{C. elegans}), fly (\emph{D.
melanogaster}), fission yeast (\emph{S. pombe}) and mouse (\emph{M.
musculus}), as well as for the human interactome \cite{Kovacs2019}.

A common way to uncover (or validate) PPIs is the \emph{Yeast-Two-Hybrid}
technique (also known as \emph{two-hybrid screening} or \emph{Y2H}),
which is based on the expression of a specific reporter gene that
activates by the binding of a DNA-binding Domain (DB) and an Activation
Domain (AD) of a Transcription Factor that binds to an Upstream Activation
Sequence (UAS). For the Y2H technique, a protein is fused to the DB
domain (known as \emph{bait}) and another one to the AD (known as
\emph{prey}). If the proteins do not interact, then the reporter gene
is not expressed. Otherwise, the reporter gene expression is activated
by the activation domain.

\begin{comment}
NL: I will make this figure myself

JF: Good idea
\end{comment}

\begin{figure}[h]
\caption{Yeast-2-Hybrid Technique}

\noindent \centering{}\includegraphics[width=1\columnwidth]{Y2H}
\end{figure}

Based on the outcome of numerous Y2H experiments PPI networks can
be constructed, in which known interactions for each protein are represented.
Several algorithms are proposed over these networks in order to predict
yet unknown interactions. The focus of this study is to evaluate different
methods for predicting PPIs using the existing knowledge of the network
of interactions, which is represented as an undirected graph. The
traditional way of approaching the problem is usually based on social
networks analysis, more specifically on the triadic closure principle
(TCP), that states that the more common shared friends that two people
have, the more likely that they know each other. As shown by previous
studies, the mentioned approach fails because it does not consider
the structural and chemical properties of the proteins \cite{Kovacs2019}.

\begin{comment}
JF: Yet you do more than evaluating this three measures so this has
to be updated.

NL: Done!

JF: Not sure we should use L3 above since it is being redefined as
a normalized measure here!
\end{comment}

For achieving the described results, human and rice PPI networks are
used and compared using state-of-the-art methods, as well as the proposed
ones (CN, A3, L3). In the case of the human network, human interactome
(\emph{HI-II-14}), as well as a curated version of it (\emph{HI-TESTED})
were used. A massive experimental assay was carried on by other authors
and its results were consolidated and used to build a validation network
(\emph{HI-III}).

\begin{comment}
JF: read throughout the document but edited up here. I prefer to read
the results before editing the materials and methods section. 
\end{comment}

\section{Materials and Methods}

\subsection{Data Availability}

Human interactome data and base source code were downloaded from the
repository of the length-3 degree normalized paths methodology \cite{Kovacs2019}:
the dataset \emph{HI-II-14} and \emph{HI-TESTED} are used for prediction
and the dataset \emph{HI-III} is used for validation.

Rice interactome information was downloaded from the STRING database
\cite{Szklarczyk2019}, corresponding to the \emph{Oryza sativa} subspecies.
The downloaded file was \emph{4530.protein.links.detailed.v11.0.txt}.
and contains more than 8 million PPIs from several resources. For
the purpose of this study and based on previous work, only PPIs with
evidence from curated databases were used (i.e. rows where the column
\emph{databases} has a value greater than zero), resulting in a network
with $5025$ nodes and $164420$ edges.

\subsection{Code Implementation}

Previous code implementation was adapted from \texttt{C++} to \texttt{Python}
(V3.6), in order to unify the algorithms into one single script. For
the purpose of algorithmic validation, the three methods were implemented
from scratch with basic functionalities and data structures of the
\texttt{Python} language.

\subsection{Data Preprocessing}

Information for the human interactome was used as-is, which corresponds
to networks of $4298$, $3727$ and $5604$ proteins and $13868$,
$9433$ and $23322$ interactions, respectively.

For the rice interactome, an additional preprocessing was performed.
The filtered network for rice consists of $5025$ proteins (nodes)
and $164420$ interactions (edges) distributed among $178$ connected
components. The connected component with the greatest number of edges
was selected in this case. The extracted connected component consists
of $n=4390$ nodes and $m=163319$ edges, which corresponds to $99.33%\%
$ of filtered edges. Further investigation is applied to this network,
which is very similar in number of nodes to the curated information
on the human interactome, although rice network is much more connected.

\subsection{Edge Prediction}

For the interaction prediction for each network, the algorithms described
below were used. It is important to keep in mind how the protein-protein
interaction (PPI) network $G=(V,E)$ is conceptualized: each node
($v_{i}\in V$) represents a protein and each undirected edge ($e_{b}=\{v_{i},v_{j}\},\,e_{b}\in E$)
represents and interaction among proteins $v_{i}$ and $v_{j}$.
\begin{description}
\item [{Common~Neighbors~(CN)}] This method is based on the Triadic Closure
Principle: ``the more common friends two individuals have, the more
likely that they know each other''. For the implementation of this
method, $A{{}^2}$ matrix is calculated, being $A$ the adjacency
matrix of the network.
\item [{Length-3~Path~Counts~(A3)}] This is the simplest implementation
of the proposed insight of ``if my friends and your friends interact,
then we might interact too''. The calculating is carried on with
$A{{}^3}$, i.e, the third power of the adjacency matrix.
\item [{Degree-normalized~L-3~Score~(L3)}] The previous approach might
overestimate the importance of some edges due to intermediate hubs
which add many shortcuts in the graph. To address that issue, a degree
normalization for the path $X\rightarrow U\rightarrow V\rightarrow Y$
is applied by considering the degree $k$ of the intermediate nodes
$U$ and $V$, as follows.
\[
p_{XY}=\sum_{U,V}\frac{A_{XU}\cdot A_{UV}\cdot A_{VY}}{\sqrt{k_{U}\cdot k_{V}}}
\]
\\
where $A_{ij}$ represents the value of the adjacency matrix for nodes
$i$ and $j$: 1 if the edge $\{i,j\}$ exists, 0 otherwise.
\end{description}

\subsection{Sampling Procedure}

For each network of protein interactions, the following procedure
was performed 10 times in order to address the stochastic nature of
the process and have a consensus:
\begin{itemize}
\item A percentage of interactions is removed at random from the network
(20\%).
\item The same amount of removed interactions are then predicted using the
main methods for prediction mentioned by Kovacs et al (2019): Common
Neighbors (\textbf{A2}), raw path count of paths of length 3 (\textbf{A3})
and the Length-3 degree-normalized score (\textbf{L3}).
\item A test dataset is created as follows: all removed edges are included
(as observed positives for the ML algorithm) and from the predicted
edges of \textbf{A2}, \textbf{A3} and \textbf{L3} that don't lie in
the previous classification (observed negatives), a random subset
is chosen such that the dataset is balanced, that is, the amount of
observed positive labels is equal to the observed negative labels.
\item Once the dataset is ready, it is randomly partitioned: 80\% is used
for \texttt{XGBoost} model training and 20\% is used for validation.
It is important to have in mind that balanced distribution of the
positive and negative labels in the datasets was satisfied in training,
but validation is performed using the remaining unbalanced data.
\end{itemize}
It is worth mentioning that some exploration experiments were carried
out with interactions removal percentages of 2, 5, 10 and 20\%, as
well as train-test partitions of 15-85, 80-20 and 75-25, and the explained
parameter set was selected because it either represented a marginal
gain (results not shown) or because it was a common parameter selection
in related literature.

\subsection{Feature Extraction with \texttt{Node2Vec}}

The \texttt{Node2Vec} module was used for extracting features of the
rice interactome graph. The parameters and considerations for the
model were:
\begin{itemize}
\item All paths in the random walks are equally likely (\texttt{p=1, q=1})
\item Use a modest number of dimensions and threads for calculation (\texttt{dimensions=16,
workers=4})
\item Since length-3 paths are the defining property in this study, there
is no necessity for longer walks. However, it is important to try
out many possible redundant routes and to consider a window of at
least 4 (\texttt{walk\_length=5, num\_walks=300, window=5})
\item Other standard parameters were left with default values (\texttt{min\_count=1,
batch\_words=4})
\item Edge embeddings were calculated using a geometric ratio of the node
embeddings (\texttt{HadamardEmbedder})
\end{itemize}

\subsection{Handcrafted Feature}

Due to the poor results of the \emph{raw} Length-3 counting (\textbf{A3}),
a different approach for this information was carried out in the present
study: As it still gives a lot of information that might be useful
for a predictive routine, this counting was normalized (dividing by
the greatest counting in the \textbf{A3} top predictions) and then
used as a feature for the Machine Learning algorithm. For completeness,
also \textbf{CN} and \textbf{L3} information was used as a possible
feature. Finally, the case were no handcrafted feature was also considered,
that is, only the features extracted from the structure of the network.

\subsection{Feature to Predict: Existence}

The feature to predict corresponds to the possible existence (\emph{True/False})
of a link based on the existing information of the network, using
the network itself in a random sub\_exploration (\texttt{Node2Vec})
as well as in a structured search (A3). This property is evaluated
by taking out a fraction of the edges and then trying to predict for
a given set of possible edges if they have a high probability to belong
to the original network.

\subsection{Machine Learning Algorithm}

The Extreme Gradient Boosting implementation of gradient boosted trees
is applied in this study to evaluate the existence of an edge. Gradient
boosted trees are usually used for supervised learning problems, where
the training data $X_{i}$ has multiple features and pretends to explain
(or predict) a target variable $Y_{i}$. The corresponding implementation
applied for this study is \texttt{XGBoost}, available publicly.

The selected parameters for the model were:\texttt{ max\_depth=3},
\texttt{colsample\_bytree=0.6} and \texttt{eval\_metric=''auc''}.

\subsection{Result Validation}

As mentioned before, 80\% of the final dataset was randomly selected
and used for training, while the remaining information was used for
validation. The whole training-validation procedure was applied 10
times.

The chosen metric for validation was the Area under the Curve (\textbf{AUC})
of the Receiver Operating Characteristic (\textbf{ROC}). This curve
corresponds to plot the sensitivity (probability of predicting a real
positive as positive) against 1-specificity (probability of predicting
a real negative as positive). It is worth to remind that AUC values
move in the range $[0,1]$, where 1 is a perfect prediction and 0.5
corresponds to a random guess. Normally, values over 0.8 of AUC are
considered good.

\section{Results and Discussion}

\subsection{Rice Interactome}

For the rice interactome, the different model-features combinations
were trained and validated. After executing the mentioned routines,
the results are shown in Figure \ref{F1}. First, one should have
a baseline of comparison, which in this case corresponds to \texttt{Node2Vec}
without any additional feature included. The plot below shows those
results, and one can see that its mean performance using the AUC metric
is $0.93$, and that the results among the 10 repetitions are consistent.
This results mean that the model using only the default features performs
very good. This result can also be assessed when looking at the confusion
matrix, where a precision of $0.9814$ and a recall of $0.8649$ can
be derived.

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{F1}Summary ROC curves for \texttt{Node2Vec} model alone}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{Only_ML/ROConlyML_SUMMARY}\includegraphics[width=0.48\columnwidth]{Only_ML/CMonlyML_SUMMARY}
\end{figure}

Next experiment carried out was to add each prediction method score
as a prediction feature, that is, use the score as the eleventh input
feature for the ML algorithm. Results for the Common Neighbors method
(\textbf{CN}), which uses paths of length 2, are shown in Figure \ref{F2}.
It can be observed again that all 10 experiments have small variability
among them and perform better than the baseline, with an area under
the ROC curve of $0.98$, with small standard deviation. When looking
at the confusion matrix for this model, turns out that most of the
guesses are true positives and true negatives, resulting in a precision
of $0.9952$ and a recall of $0.9108$. Same analyses are done with
the count of paths of length 3 (\textbf{A3}) and with the degree-normalized
length-3 score (\textbf{L3}) and results are presented in figures
\ref{F3} and \ref{F4}, respectively, resulting also in area under
the curve of $0.97$ and $0.98$, respectively.

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{F2}Results for \texttt{Node2Vec} with CN feature}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{ML_Metric/ROCA2_SUMMARY}\includegraphics[width=0.48\columnwidth]{ML_Metric/CMA2_SUMMARY}
\end{figure}

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{F3}Results for \texttt{Node2Vec} with A3 feature}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{ML_Metric/ROCA3_SUMMARY}\includegraphics[width=0.48\columnwidth]{ML_Metric/CMA3_SUMMARY}
\end{figure}

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{F4}Results for \texttt{Node2Vec} with L3 feature}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{ML_Metric/ROCL3_SUMMARY}\includegraphics[width=0.48\columnwidth]{ML_Metric/CML3_SUMMARY}
\end{figure}

The precision when using the count of paths of length 3 (A3) is $0.9935$
while the recall is $0.9055$. It is interesting to observe that precision
decreased $0.16\%$ and recall decreased $0.58\%$. A similar situation
is obtained for the model with the normalized score (L3), whose precision
decreased to $0.9944$ ($-0.08\%$) and recall decreased to $0.8891$
($-2.38\%$). It can be seen also that the ROC curve achieves high
values of true positive rate sooner in the A2-featured model than
on the other models.

Another relevant assessment was carried out to verify the predictive
power of the metrics. \texttt{XGBoost} models were trained only using
each metric, for the same 10 experiments. Results are shown in Figures
\ref{F5}, \ref{F6} and \ref{F7}. It can be observed that XGBoost
performs worse when only considering CN, A3 and L3 alone, although
good values of precision (CN:0.9916, A2:0.9807 and A3:0.9838) and
recall (CN:0.6704, A2:0.7012 and A3:0.7037) are achieved.

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{F5}Results for CN feature only}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{Only_Metric/ROConlyA2_SUMMARY}\includegraphics[width=0.48\columnwidth]{Only_Metric/CMonlyA2_SUMMARY}
\end{figure}

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{F6}Results for A3 feature alone}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{Only_Metric/ROConlyA3_SUMMARY}\includegraphics[width=0.48\columnwidth]{Only_Metric/CMonlyA3_SUMMARY}
\end{figure}

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{F7}Results for L3 feature alone}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{Only_Metric/ROConlyL3_SUMMARY}\includegraphics[width=0.48\columnwidth]{Only_Metric/CMonlyL3_SUMMARY}
\end{figure}

Finally, an relevant evaluation on the models that include a handcrafted
feature is necessary: How important is the appended feature for the
model result? The answer can be observed by the feature importance
plots for CN, A3 and L3 in Figure \ref{F8-importance}. Each plot
resembles the importance gain from each feature in the underlying
decision trees that \texttt{XGBoost} uses. Usually, the more bifurcations,
the higher gain a feature has, the more relevance a feature has on
the model prediction itself. All models have the appended feature
as the most important by a good margin, although margin for A3 is
not wide as for A2 or L3.

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{F8-importance}Importance gain plots for \texttt{Node2Vec}
with each feature}
\par\end{centering}
\begin{centering}
\includegraphics[width=0.7\columnwidth]{ML_Metric/Imp\lyxdot A2\lyxdot All}
\par\end{centering}
\begin{centering}
\includegraphics[width=0.7\columnwidth]{ML_Metric/Imp\lyxdot A3\lyxdot All}
\par\end{centering}
\centering{}\includegraphics[width=0.7\columnwidth]{ML_Metric/Imp\lyxdot L3\lyxdot All}
\end{figure}

\subsection{Human Interactome}

As the first step in the evaluation of CN, A3 and L3 methods for edge
prediction, the full calculation and ranking of all predictions was
carried out. The results for the analyzed human interactomes (\emph{HI-II-14}
and \emph{HI-TESTED}) is shown in Figure \ref{fig:HI1}, in which
the x axis represents a rank position $k$ and the y axis displays
the precision for the top $k$ predictions of each method, when assuming
interactome \emph{HI-III} as the validation set.

\begin{figure}[h]
\caption{\label{fig:HI1}Methods Comparison for \emph{HI-II-14} and \emph{HI-TESTED}}

\noindent \centering{}\includegraphics[width=0.48\columnwidth]{hi-ii-14\lyxdot txt}\includegraphics[width=0.48\columnwidth]{hi-tested\lyxdot txt}
\end{figure}

As it can be inferred from the plots, L3-based predictions outperform
their $A{{}^2}$ counterparts. Results also show that L3-score and
$A^{3}$predictions follow a very similar trend.

However, it is of our interest to assess if Machine Learning can boost
the overall performance on the prediction, so the proposed procedure
was adapted to the human interactome data: Instead of having to remove
a fraction of edges from the network in order to predict and validate,
now the prediction network and the validation networks are given beforehand.
The rest of the procedure is carried out as mentioned before. Results
for the Interactome \emph{HI-II-14} with Node2Vec and each metric
are presented in Figures \ref{HI1}, \ref{HI2} and \ref{HI3}.

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{HI1}\emph{HI-II-14} results for \texttt{Node2Vec} with CN
feature}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{Human/ROC\lyxdot A2}\includegraphics[width=0.48\columnwidth]{Human/CM\lyxdot A2}
\end{figure}

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{HI2}\emph{HI-II-14} results for \texttt{Node2Vec} with A3
feature}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{Human/ROC\lyxdot A3}\includegraphics[width=0.48\columnwidth]{Human/CM\lyxdot A3}
\end{figure}

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{HI3}\emph{HI-II-14} results for \texttt{Node2Vec} with L3
feature}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{Human/ROC\lyxdot L3}\includegraphics[width=0.48\columnwidth]{Human/CM\lyxdot L3}
\end{figure}

It can be seen that for \emph{HI-II-14}, CN performs marginally better
than L3 when comparing precision ($0.9978$ vs $0.9977$) but worse
in terms of recall ($0.9346$ vs $0.9598$), although all metrics
have a good values for area under the ROC curve of 0.99.

On the other hand, the interactome \emph{HI-TESTED} was assessed with
the same methodology as \emph{HI-II-14} and results are shown in Figures
\ref{HI1-1}, \ref{HI2-1} and \ref{HI3-1}. It can be seen that in
terms of false positives, CN performs better than A3 and L3, but in
terms of false negatives, A3 actually performs better that CN and
L3. In terms of AUC, metrics results are indistinguishable.

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{HI1-1}\emph{HI-TESTED} results for \texttt{Node2Vec} with
CN feature}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{Human/ROC\lyxdot A2\lyxdot 2}\includegraphics[width=0.48\columnwidth]{Human/CM\lyxdot A2\lyxdot 2}
\end{figure}

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{HI2-1}\emph{HI-TESTED} results for \texttt{Node2Vec} with
A3 feature}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{Human/ROC\lyxdot A3\lyxdot 2}\includegraphics[width=0.48\columnwidth]{Human/CM\lyxdot A3\lyxdot 2}
\end{figure}

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{HI3-1}\emph{HI-TESTED} results for \texttt{Node2Vec} with
L3 feature}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{Human/ROC\lyxdot L3\lyxdot 2}\includegraphics[width=0.48\columnwidth]{Human/CM\lyxdot L3\lyxdot 2}
\end{figure}

An interesting point from this human datasets can be observed when
looking at the importance plots, that show a very biased influence
of the handcrafted features. Figure X shows this behavior for the
\emph{HI-II-14} dataset, which is very similar to \emph{HI-TESTED}.

\begin{figure}[H]
\noindent \begin{centering}
\caption{\label{F8-importance-1}Importance gain plots for \emph{HI-II-14}}
\par\end{centering}
\begin{centering}
\includegraphics[width=0.7\columnwidth]{Human/Imp\lyxdot A2\lyxdot 1}
\par\end{centering}
\begin{centering}
\includegraphics[width=0.7\columnwidth]{Human/Imp\lyxdot A3\lyxdot 1}
\par\end{centering}
\centering{}\includegraphics[width=0.7\columnwidth]{Human/Imp\lyxdot L3\lyxdot 1}
\end{figure}

\section{Conclusions}

Based on previous work from other authors\cite{Kovacs2019}, the intrinsic
knowledge of existing protein-protein interaction networks let scientists
elucidate promising new interactions to explore. In this study, a
machine learning algorithm was fed with previously elucidated information
about the network, as well as results form a state-of-the-art edge
embedding technique. Even if the baseline of the handcrafted features
(CN, A3, L3) yields good results, boosting that information with the
vector embedding, which calculates vector representations of the network
structure improves the overall performance of each method on its own.

It was found that applying \texttt{XGBoost} to the rice PPI prediction
obtained very good results when only using the node embedding technique
(AUC=0.93) and when only using the handcrafted features based on protein
neighborhood (AUC \textasciitilde 0.88). Even better results were
achieved when the handcrafted features and the vector representation
information were combined (AUC \textasciitilde 0.98). It is important
to consider that the importance gain of the handcrafted feature in
each model was high, although importance values for the A3 method
stand out less from the node embedding information than in CN and
L3.

On the human interactome, the results show that on both interactomes
(\emph{HI-II-14} and \emph{HI-TESTED}) all metrics yield very good
results for the link prediction problem, and therefore this study
extends previous work by other authors that only consider the predictive
power of the metrics alone.

It is important to note the difference in feature importance from
rice and human interactomes. In the former, features have a more balanced
importance distribution than in the latter dataset, whose model shows
a very strong dependence on the handcrafted feature.

With this study, a framework for link prediction on protein-protein
interactions is presented and validated against the available information
on rice and human interactomes, but can be easily extended to other
organisms or even to other graphs not necessarily related to biology.

\bibliographystyle{plain}
\bibliography{refs}

\end{document}