### 2020-12-07 camila/report/declarations.tex

\section*{Acknowledgements}%% if any
Not applicable

\section*{Funding}%% if any
This work was funded by the OMICAS program: Optimización Multiescala In-silico de
Cultivos Agrícolas Sostenibles (Infraestructura y Validación en Arroz y Caña de Azúcar),
anchored at the Pontificia Universidad Javeriana in Cali and funded within the Colombian
Scientific Ecosystem by The World Bank, the Colombian Ministry of Science, Technology and
Innovation, the Colombian Ministry of Education and the Colombian Ministry of Industry 
and Turism, and ICETEX, under GRANT ID: FP44842-217-2018.

\section*{Abbreviations}%% if any
\textbf{HLC:} Hierarchical Link Clustering\\
\textbf{LASSO:} Least Absolute Shrinkage Selector Operator\\
\textbf{PCC:} Pearson Correlation Coefficient\\
\textbf{RNA-seq:} RNA sequencing\\
\textbf{SHT2:} Spermidine hydroxycinnamoyltransferase 2\\
\textbf{WGCNA:} Weighted Gene Co-expression Network Analysis

\section*{Availability of data and materials}%% if any
The datasets analysed during the current study are publicy available. They can be found in the following locations:
\begin{itemize}
\item RNA-seq data of salt stress in rice is available on the GEO (GSE98455).
\item Phenotypic data of salt stress in rice is a subset of the supplementary file 1 included in ~\cite{campbell2017allelic}.
\end{itemize}

\section*{Ethics approval and consent to participate}%% if any
Not applicable

\section*{Competing interests}
The authors declare that they have no competing interests.

\section*{Consent for publication}%% if any
Not applicable

\section*{Authors' contributions}
J.F. and C.R. proposed the original idea. 
J.F. provide advice on algorithms concepts and implementation.
C.R. structured the methodology and performed the analysis. 
C.R., J.F., and H.C.R. wrote the manuscript.
All authors read and approved the final manuscript.