### 2020-12-11 nicolas/PPIPL3/bmc_template/declarations.tex

\section*{Acknowledgements}%% if any
N.L. would like to thank Camila Riccio and Miguel Romero-Diaz for their help 
during the research.

\section*{Funding}%% if any
This work was funded by the OMICAS program: Optimización Multiescala In-silico de
Cultivos Agrícolas Sostenibles (Infraestructura y Validación en Arroz y Caña de Azúcar),
anchored at the Pontificia Universidad Javeriana in Cali and funded within the Colombian
Scientific Ecosystem by The World Bank, the Colombian Ministry of Science, Technology and
Innovation, the Colombian Ministry of Education and the Colombian Ministry of Industry 
and Turism, and ICETEX, under GRANT ID: FP44842-217-2018.

\section*{Abbreviations}%% if any
\textbf{PPI:} Protein-Protein Interaction\\
\textbf{Y2H:} Yeast-2-Hybrid screening\\
\textbf{AUC:} Area under the curve\\
\textbf{ROC:} Receiver Operating Characteristic\\
\textbf{FP:} False Positive\\
\textbf{FN:} False Negative\\
\textbf{TP:} True Positive\\
\textbf{TN:} True Negative

\section*{Availability of data and materials}%% if any
The datasets analysed during the current study are publicy available. They can be found in the following locations:
\begin{itemize}
\item Human interactome information was retrieved of the supplementary files included in ~\cite{Kovacs2019}.
\item Rice Interactomes were downloaded from STRING database of protein interactions (4530.protein.links.detailed.v11.0.txt).
Only interactions which have a curated origin (meaning a value different than zero on the column \emph{'databases'})
\end{itemize}

\section*{Ethics approval and consent to participate}%% if any
Not applicable

\section*{Competing interests}
The authors declare that they have no competing interests.

\section*{Consent for publication}%% if any
Not applicable

\section*{Authors' contributions}
N.L. proposed the original idea. 
J.F. provide advice on algorithms, libraries and implementation.
N.L. performed the implementation and analysis. 
N.L., J.F., and C.R. wrote the manuscript.
All authors read and approved the final manuscript.