### 2020-12-28 nicolas/PPIPL3/bmc_template/case.tex

% case.tex

For the rice interactome, different combinations of features were
trained on the model and validated. The results are shown in Table
\ref{T2}. First, prediction based on \texttt{Node2Vec} without any
additional feature included is used as baseline of comparison. Figure
\ref{F1} shows those results, and one can see that its mean performance
using the AUC is $0.93$. The results for all 10 repetitions of each
model are consistent, meaning the variance among each line is not
significant. The results for each model using only the default features
are satisfactory. This result can also be assessed when looking at
the confusion matrix, where a precision of $0.98$ and a recall of
$0.86$ are obtained.

\begin{table}
\caption{\label{T2}Summary statistics for rice interactome}
\includegraphics[width=1\columnwidth]{figures/T2.eps}
\end{table}

\begin{figure}[h]
\noindent \begin{centering}
\caption{\label{F1}Summary ROC curves for \texttt{Node2Vec} model alone}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{figures/Only_ML/ROConlyML_SUMMARY.eps}\includegraphics[width=0.48\columnwidth]{figures/Only_ML/CMonlyML_SUMMARY.eps}
\end{figure}

The next experiment carried out was to add for each prediction method
score as a prediction feature, that is, use the score as an additional
feature for the ML algorithm. Results for the Common Neighbors method
(\textbf{A2}), which uses paths of length 2, are shown in Figure S\ref{F2}.
All 10 experiments have small variability among them and perform better
than the baseline, with AUC of $0.98$. When looking at the confusion
matrix for this model, most guesses are true positives and true negatives,
resulting in a precision of $0.99$ and a recall of $0.91$. The same
analyses are done with the count of paths of length 3 (\textbf{A3})
and with the degree-normalized length-3 score (\textbf{L3}). These
results are presented in figures S\ref{F3} and \ref{F4}, respectively.
The corresponding AUC are $0.97$ and $0.98$.

\begin{figure}[h]
\noindent \begin{centering}
\caption{\label{F4}Results for \texttt{Node2Vec} with L3 feature}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{figures/ML_Metric/ROCL3_SUMMARY.eps}\includegraphics[width=0.48\columnwidth]{figures/ML_Metric/CML3_SUMMARY.eps}
\end{figure}

The precision when using the count of paths of length 3 (A3) is $0.99$
while the recall is $0.90$. It is interesting to observe that precision
decreased $0.16\%$ and recall decreased $0.58\%$. A similar situation
is obtained for the model with the normalized score (L3), whose precision
decreased to $0.99$ ($-0.08\%$) and recall decreased to $0.88$
($-2.38\%$). It can be seen also that the ROC curve achieves high
values of true positive rate sooner in the A2-featured model than
on the other models.

Another relevant assessment was carried out to verify the predictive
power of the metrics. \texttt{XGBoost} models were trained only using
each metric, for the same 10 experiments. Results are shown in Figures
S\ref{F5}, S\ref{F6} and \ref{F7}. XGBoost yields lower AUC values
when only considering A2, A3 and L3. Precision and recall for this
assessments, although lower than the previous experiments, are still
considered satisfactory.

\begin{figure}[h]
\noindent \begin{centering}
\caption{\label{F7}Results for L3 feature alone}
\par\end{centering}
\noindent \raggedleft{}\includegraphics[width=0.48\columnwidth]{figures/Only_Metric/ROConlyL3_SUMMARY.eps}\includegraphics[width=0.48\columnwidth]{figures/Only_Metric/CMonlyL3_SUMMARY.eps}
\end{figure}

Finally, an relevant evaluation on the models that include a handcrafted
feature is necessary: How important is the appended feature for the
model result? The answer can be observed by the feature importance
plots for A2, A3 and L3 in Figure \ref{F8-importance}. Each plot
resembles the importance gain from each feature in the underlying
decision trees that \texttt{XGBoost} uses. Usually, the more bifurcations,
the higher gain a feature has, the more relevance a feature has on
the model prediction itself. All models have the appended feature
as the most important by a good margin, although margin for A3 is
not wide as for A2 or L3.

\begin{figure}[h]
\noindent \begin{centering}
\caption{\label{F8-importance}Importance gain plots for \texttt{Node2Vec}
with each feature}
\par\end{centering}
\begin{centering}
\includegraphics[width=0.48\columnwidth]{figures/ML_Metric/Imp\lyxdot A2\lyxdot All.eps}\includegraphics[width=0.48\columnwidth]{figures/ML_Metric/Imp\lyxdot A3\lyxdot All.eps}
\par\end{centering}
\centering{}\includegraphics[width=0.48\columnwidth]{figures/ML_Metric/Imp\lyxdot L3\lyxdot All.eps}
\end{figure}